Non-modal Growth in LAPD Turbulence

### 2013-11-08 Transient_growth_prl.tex
Nov 8

@@

In these equations, $\mu_N$, $\mu_T$, and $\mu_\phi$ are artificial diffusion and viscosity coefficients used for subgrid dissipation. They are large enough to allow saturation
and grid convergence~\cite{friedman2012}, but small enough to allow for turbulence to develop. In the simulations, they are all given the same value of $1.25 \times 10^{-3}$ in Bohm-normalized units. 
This is the only free parameter in the simulations. All other parameters such as the electron collisionality $\nue$, ion-neutral
collisionality $\nuin$, parallel electron thermal conductivity $\kpe$, and mass ratio $\fmie$ are calculated from experimental quantities.
There are two sources of free energy: the density gradient due to the equilibrium density profile $N_0$, and the equilibrium electron temperature gradient in $T_{e0}$, both of which are
taken from experimental fits. $N_0$ and $T_{e0}$ are functions of only the radial cylindrical coordinate $r$, and they are shown in Fig.~\ref{eq_profiles}. 
The mean potential profile $\phi_0$ is set to zero in the model, and terms involving $\phi_0$ are not included in Eqs.~\ref{ni_eq}-\ref{te_eq}. 
The justification for this is that biasable azimuthal limiters in LAPD allow for the mean $E \times B$ flow and flow shear to be varied with high precision, even allowing the flow to be
nulled out~\cite{schaffner2012}. 
The simulations in this paper use the $N_0$, $T_{e0}$, and $\phi_0$ profiles from the nulled out flow experiment, justifying setting the mean potential profile to zero in the simulations.

>>>>

Since the early 1990's, it has become clear that in dynamical systems with energetically conservative nonlinearities, 
the linear operator must be non-normal for subcritical turbulence to occur~\cite{trefethen1993}.
Non-normality is the condition that an operator or matrix does not commute with its adjoint. One consequence of non-normality is that the linear eigenvectors of the operator are nonorthogonal.
In non-normal systems, even when all of the eigenvalues lie in the stable domain (where all linear eigenvectors decay exponential), 
certain types of fluctuations in a system can access the free energy in the equilibrium gradients and transiently amplify the fluctuation energy in the system. 
A paradigmatic illustration of this transient growth process may be seen if Figure 2 of a review by Schmid~\cite{schmid2007}.
Such behavior is obscured by traditional normal mode analysis, which in non-normal systems, only effectively describes the long time asymptotic behavior of fluctuations under the 
action of the linear operator. Transient events, such as those that can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
Non-modal analysis, in fact, has provided the tool that can explain and predict subcritical turbulence.

While the hydrodynamics community has embraced non-modal analysis and transient growth over the past two decades, the plasma community
still heavily relies on normal mode analysis to inform turbulent predictions and observations, often without checking for non-modal effects. 
While this has been successful in certain facets of plasma physics, it has failed in others.

Direct numerical simulations of plasma turbulence have become increasingly common with the increase in computing power. These simulations produce so much data that the results can be difficult to
comprehend. It is often easiest to compare simulation results with simple models to achieve some understanding.

@@

In these equations, $\mu_N$, $\mu_T$, and $\mu_\phi$ are artificial diffusion and viscosity coefficients used for subgrid dissipation. They are large enough to allow saturation
and grid convergence~\cite{friedman2012}, but small enough to allow for turbulence to develop. In the simulations, they are all given the same value of $1.25 \times 10^{-3}$ in Bohm-normalized units. 
This is the only free parameter in the simulations. All other parameters such as the electron collisionality $\nue$, ion-neutral
collisionality $\nuin$, parallel electron thermal conductivity $\kpe$, and mass ratio $\fmie$ are calculated from experimental quantities.
There are two sources of free energy: the radial density gradient due to the equilibrium density profile $N_0$, and the equilibrium radial electron temperature gradient in $T_{e0}$, both of which are
taken from experimental fits. 
The mean potential profile $\phi_0$ is set to zero in the model, and terms involving $\phi_0$ are not included in Eqs.~\ref{ni_eq}-\ref{te_eq}. 
The justification for this is that biasable azimuthal limiters in LAPD allow for the mean $E \times B$ flow and flow shear to be varied with high precision, even allowing the flow to be
nulled out~\cite{schaffner2012}. 
The simulations in this paper use the $N_0$, $T_{e0}$, and $\phi_0$ profiles from the nulled out flow experiment, justifying setting the mean potential profile to zero in the simulations.
